---
title: "Public flow cytometry data exploration"
output: html_notebook
---

# Setup
```{r}
# # Install 'flowCore' package
# if (!require("BiocManager",
#              quietly = TRUE))
#   install.packages("BiocManager")
# BiocManager::install("flowCore")
# BiocManager::install("modreg")

require(flowCore)
require(data.table)
require(ggplot2)

# DNAMR Dependencies
require(class)
require(MASS)
require(nnet)
require(e1071)
require(moments)
require(modreg)

require(DNAMR)
```

Fluorescent Dye Directory:  
https://app.fluorofinder.com/dyes?gclid=Cj0KCQjwyt-ZBhCNARIsAKH1176G3SpZ4jKJg3nIZvZKAzYVae4ToSHajbJNHrPKx6usUlVIhUDt3bAaAkBmEALw_wcB  
  
Commonly Used Fluorochromes and Their Properties  
https://www.umc.edu/cancerinstitute/Cancer-Research/Core-Facilities/Flow%20Cytometry%20Core/Resources/Commonly-Used-Fluorochromes-and-Properties.html 

Quality control with flowAI:  
https://bioconductor.org/packages/release/bioc/html/flowAI.html  
https://bioconductor.org/packages/release/bioc/vignettes/flowAI/inst/doc/flowAI.html  
  
# All Data
Data Repository: https://flowrepository.org/public_experiment_representations 

#  Data: FR-FCM-ZZZU (20190311 TOPRO GFP BFP mCherry PECy7 BV510)
## Load data
Cell isolation by FACS of lncRNAs targeted with pgRNAs along BLaER1 transdifferentiation.  
Source: https://flowrepository.org/experiments/3892/download_ziped_files  
  
```{r}
fname <- list.files(path = "Data/FlowRepository_FR-FCM-Z3PK_files",
                    pattern = "\\.fcs$")
fname

l1 <- lapply(X = fname,
             FUN = function(a) {
               tmp <- read.FCS(paste0("Data/FlowRepository_FR-FCM-Z3PK_files/",
                                      a))
               return(data.table(`File Name` = a,
                                 tmp@exprs))
             })
names(l1) <- fname

head(l1[[1]])
dim(l1[[1]])
names(l1[[1]])
```

## Meta data
Source: https://flowrepository.org/experiments/3892  
  
```{r}
# Sample info
fread("Data/FlowRepository_FR-FCM-Z3PK_files/meta.csv",
      header = FALSE,
      nrows = 6)

# Meta data
meta <- fread("Data/FlowRepository_FR-FCM-Z3PK_files/meta.csv",
              header = TRUE,
              skip = 6)
```

## Long data
```{r}
dtl <- rbindlist(l1)
dtl <- merge(meta,
             dtl,
             by = "File Name")
dtl$`File Name` <- factor(dtl$`File Name`,
                          levels = unique(dtl$`File Name`))

dtl$Description <- factor(dtl$Description,
                          levels = unique(dtl$Description))

dtl$`Sample characteristic` <- factor(dtl$`Sample characteristic`,
                                      levels = c("Undifferentiated",
                                                 "Differentated T3"))

dtl$`Experiment Variables` <- factor(dtl$`Experiment Variables`,
                                     levels = c("Negative control, T0",
                                                "Negative control, T3",
                                                "lncRNAs, T0",
                                                "lncRNAs, T3"))

dtl$Treatment <- factor(dtl$Treatment,
                        levels = c("Negative control",
                                   "lncRNAs"))
summary(dtl)
```

## Save data
```{r}
save(dtl,
     file = "Data/dtl.RData")
```

## START HERE: Reload data
```{r}
load("Data/dtl.RData")
```

## Forward scattering (FSC) vs. Side Scattering (SSC)
FSC-A: Forward Scattering Area  
FSC-H: Forward Scattering Height  
SSC-A: Side Scattering Area  
SSC-H: Side Scattering Height   
  
Source: [A guide to gating in flow cytometry](https://www.bio-rad-antibodies.com/blog/a-guide-to-gating-in-flow-cytometry.html)

## Singletons vs. cell clusters
### Raw scale
```{r,fig.width=6,fig.height=7}
ggplot(dtl,
       aes(x = `FSC-A`,
           y = `FSC-H`,
           color = `Experiment Variables`)) +
  facet_wrap(~ `Sample characteristic` + Treatment) +
  geom_point(shape = ".",
             alpha = 0.3) +
  # geom_density_2d() +
  stat_density_2d(contour_var = "count",
                  color = "black") +
  theme_bw() +
  theme(legend.position = "none")
```

## Log-log scale
```{r,fig.width=6,fig.height=7}
ggplot(dtl,
       aes(x = log(`FSC-A` - min(`FSC-A`) + 10000),
           y = log(`FSC-H` - min(`FSC-H`) + 10000),
           color = `Experiment Variables`)) +
  facet_wrap(~ `Sample characteristic` + Treatment) +
  geom_point(shape = ".",
             alpha = 0.3) +
  stat_density_2d(contour_var = "count",
                  color = "black") +
  geom_abline(slope = 1,
              intercept = 0,
              linetype = "dashed") +
  theme_bw() +
  theme(legend.position = "none")
```

## Javier's transformation
```{r}
Z <- cbind(A = log(dtl$`FSC-A` - min(dtl$`FSC-A`) + 10000),
           H = log(dtl$`FSC-H` - min(dtl$`FSC-H`) + 10000))
s <- var(Z)
s

xx <- chol(solve(s))
xx

xxx <- Z %*% t(xx)
plot(xxx, 
     pch = ".",
     col = alpha(3, 0.3))
abline(v = -7,
       col = 2,
       lty = 2)
abline(h = 27,
       col = 2,
       lty = 2)
```

## Remove cell clusters
```{r}
ndx <- which(!(xxx[, 1] > -7 &
               xxx[, 2] > 27))

dt2 <- dtl[ndx, ]
```

# Debris
## Areas, raw scale
```{r,fig.width=6,fig.height=7}
ggplot(dtl,
       aes(x = `FSC-A`,
           y = `SSC-A`,
           color = `Experiment Variables`)) +
  facet_wrap(~ `Sample characteristic` + Treatment) +
  geom_point(shape = ".",
             alpha = 0.3) +
  stat_density_2d(contour_var = "count",
                  color = "black") +
  theme_bw() +
  theme(legend.position = "none")

ggplot(dt2,
       aes(x = `FSC-A`,
           y = `SSC-A`,
           color = `Experiment Variables`)) +
  facet_wrap(~ `Sample characteristic` + Treatment) +
  geom_point(shape = ".",
             alpha = 0.3) +
  stat_density_2d(contour_var = "count",
                  color = "black") +
  theme_bw() +
  theme(legend.position = "none")
```

## Areas, log scale
```{r,fig.width=6,fig.height=7}
ggplot(dtl,
       aes(x = log(`FSC-A` - min(`FSC-A`) + 10000),
           y = log(`SSC-A` - min(`SSC-A`) + 10000),
           color = `Experiment Variables`)) +
  facet_wrap(~ `Sample characteristic` + Treatment) +
  geom_point(shape = ".",
             alpha = 0.3) +
  stat_density_2d(contour_var = "count",
                  color = "black") +
  geom_abline(slope = 1,
              intercept = 0,
              linetype = "dashed") +
  theme_bw() +
  theme(legend.position = "none")

ggplot(dt2,
       aes(x = log(`FSC-A` - min(`FSC-A`) + 10000),
           y = log(`SSC-A` - min(`SSC-A`) + 10000),
           color = `Experiment Variables`)) +
  facet_wrap(~ `Sample characteristic` + Treatment) +
  geom_point(shape = ".",
             alpha = 0.3) +
  stat_density_2d(contour_var = "count",
                  color = "black") +
  geom_abline(slope = 1,
              intercept = 0,
              linetype = "dashed") +
  theme_bw() +
  theme(legend.position = "none")
```

## Height, raw scale
```{r,fig.width=6,fig.height=7}
ggplot(dtl,
       aes(x = `FSC-H`,
           y = `SSC-H`,
           color = `Experiment Variables`)) +
  facet_wrap(~ `Sample characteristic` + Treatment) +
  geom_point(shape = ".",
             alpha = 0.3) +
  stat_density_2d(contour_var = "count",
                  color = "black") +
  theme_bw() +
  theme(legend.position = "none")

ggplot(dt2,
       aes(x = `FSC-H`,
           y = `SSC-H`,
           color = `Experiment Variables`)) +
  facet_wrap(~ `Sample characteristic` + Treatment) +
  geom_point(shape = ".",
             alpha = 0.3) +
  stat_density_2d(contour_var = "count",
                  color = "black") +
  geom_vline(xintercept = 0.25*10^5,
             linetype = "dashed",
             color = "black") +
  theme_bw() +
  theme(legend.position = "none")
```

## Remove debris
```{r}
dt3 <- dt2[dt2$`FSC-H` > 0.25*10^5, ]
```


# ASIDE: Signal to Pulse to Dot
Pages 19-24 of:  
https://www.biovis.uu.se/digitalAssets/593/c_593618-l_1-k_flowbasics_biovis_pacho_190626.pdf  
  
```{r,fig.width=6,fig.height=7}
ggplot(dtl,
       aes(x = `FSC-H`,
           y = `SSC-A`,
           color = `Experiment Variables`)) +
  facet_wrap(~ `Sample characteristic` + Treatment) +
  geom_point(shape = ".",
             alpha = 0.3) +
  stat_density_2d(contour_var = "count",
                  color = "black") +
  theme_bw() +
  theme(legend.position = "none")
```

# Fluorochromes
```{r,fig.width=6,fig.height=7}
ggplot(dt3,
       aes(x = log(`Pacific Blue-A` - min(`Pacific Blue-A`) + 100),
           y = log(`PE-Texas Red-A` - min(`PE-Texas Red-A`) + 100),
           color = `Experiment Variables`)) +
  facet_wrap(~ `Sample characteristic` + Treatment) +
  geom_point(shape = ".",
             alpha = 0.3) + 
  theme_bw() +
  theme(legend.position = "none")
```

```{r,fig.width=6,fig.height=7}
ggplot(dt3,
       aes(x = log(`AmCyan-A` - min(`AmCyan-A`) + 100),
           y = log(`PE-Cy7-A` - min(`PE-Cy7-A`) + 100),
           color = `Experiment Variables`)) +
  facet_wrap(~ `Sample characteristic` + Treatment) +
  geom_point(shape = ".") +
  theme_bw() +
  theme(legend.position = "none")
```

# 2D density plot
Source: https://ggplot2.tidyverse.org/reference/geom_density_2d.html  
  
```{r,fig.width=6,fig.height=7}
ggplot(dt3,
       aes(x = log(`AmCyan-A` - min(`AmCyan-A`) + 1),
           y = log(`PE-Cy7-A` - min(`PE-Cy7-A`) + 1),
           color = `Experiment Variables`)) +
  facet_wrap(~ `Sample characteristic` + Treatment) +
  geom_point(shape = ".",
             color = "grey") +
  # geom_density_2d() +
  stat_density_2d(contour_var = "count") +
  theme_bw() +
  theme(legend.position = "none")
```

# PCA
```{r}
dt_pca <- dt3[, `Pacific Blue-A`:`QDot 800-A`]
dt_pca <- apply(X = dt_pca,
                MARGIN = 2,
                FUN = function(a) {
                  return(log(a - min(a) + 1))
                })
```

```{r}
m1 <- prcomp(dt_pca,
             center = TRUE,
             scale. = TRUE)
summary(m1)
```

## Scores, i.e. points (df.u)
```{r}
choices <- 1:2
dt.scr <- data.table(m1$x[, choices])

# Add grouping variable
dt.scr$grp <- factor(dt3$`Experiment Variables`)
head(dt.scr)
```

## Loadings, i.e. arrows (df.v)
```{r}
dt.rot <- as.data.frame(m1$rotation[, choices])
dt.rot$feat <- rownames(dt.rot)
dt.rot <- data.table(dt.rot)
dt.rot
```

## Axis labels
```{r}
u.axis.labs <- paste(colnames(dt.rot)[1:2], 
                     sprintf('(%0.1f%% explained var.)', 
                             100*m1$sdev[choices]^2/sum(m1$sdev^2)))
u.axis.labs
```

## Biplot
```{r,fig.width=7,fig.height=8}
scl <- 20

p1 <- ggplot(data = dt.rot,
             aes(x = PC1,
                 y = PC2)) +
  facet_wrap(~ grp) +
  geom_point(data = dt.scr,
             aes(color = grp),
             shape = ".",
             alpha = 0.5) +
  geom_segment(data = dt.rot,
               aes(x = 0,
                   y = 0,
                   xend = scl*PC1,
                   yend = scl*PC2),
               arrow = arrow(length = unit(1/2, 'picas')),
               color = "black",
               inherit.aes = FALSE) +
  geom_text(data = dt.rot,
            aes(x = 1.1*scl*PC1,
                y = 1.1*scl*PC2,
                label = feat),
            size = 2,
            hjust = 0.5,
            inherit.aes = FALSE) +
  scale_x_continuous(u.axis.labs[1]) +
  scale_y_continuous(u.axis.labs[2]) +
  scale_color_discrete("") +
  ggtitle("") +
  theme_bw() +
  theme(legend.position = "none")

tiff(filename = "tmp/pca_biplot.tiff",
     width = 7,
     height = 8,
     units = 'in',
     res = 300,
     compression = "lzw+p")
print(p1)
graphics.off()

p1
```

# Data: FR-FCM-ZZZU
0001.FCS - UE, unstimulated  
0008.FCS - UE, unstimulated  
0003.FCS - UE, LPS stimulated  
0010.FCS - UE, LPS stimulated  
   
UE - UN-EXPOSED infants; HUE - HIV-EXPOSED-UNINFECTED infants  
  
## Load data
```{r}
fname <- list.files(path = "Data/FlowRepository_FR-FCM-ZZZU_files/",
                    pattern = "\\.FCS$")
fname

l2 <- lapply(X = fname,
             FUN = function(a) {
               tmp <- read.FCS(paste0("Data/FlowRepository_FR-FCM-ZZZU_files/",
                                      a))
               return(data.table(`File Name` = a,
                                 tmp@exprs))
             })
names(l2) <- fname

head(l2[[1]])
dim(l2[[1]])
names(l2[[1]])
```

## Long data
0001.FCS - UE, unstimulated  
0003.FCS - UE, LPS stimulated  
0008.FCS - UE, unstimulated 
0010.FCS - UE, LPS stimulated 
  
```{r}
dtl <- rbindlist(l2)

meta <- data.table(`File Name` = names(l2),
                   Group = rep(c("UE, unstimulated",
                                 "UE, LPS stimulated"),
                               2))
meta$Group <- factor(meta$Group,
                     levels = unique(meta$Group))

dtl <- merge(meta,
             dtl,
             by = "File Name")
dtl$`File Name` <- factor(dtl$`File Name`,
                          levels = unique(meta$`File Name`))

summary(dtl)
```

## 
```{r}
dye <- data.table(dye = colnames(dtl)[5:12],
                  protein = c("IFNa",
                              "CD123",
                              "MHCII",
                              "CD14",
                              "CD11c",
                              "IL6",
                              "IL12",
                              "TNFa"))
dye
```

## Save data
```{r}
save(dtl,
     file = "Data/dtl_FR-FCM-ZZZU.RData")
gc()
```

## START HERE: Reload data
```{r}
load("Data/dtl_FR-FCM-ZZZU.RData")
```

## Forward scattering (FSC) vs. Side Scattering (SSC)
FSC-A: Forward Scattering Area  
SSC-A: Side Scattering Area  
FITC-A,  PE-A, PerCP-Cy5-5-A, PE-Cy7-A, APC-A, APC-Cy7-A,  Pacific Blue-A, Alex 700-A: florochromes   
  
Source: [A guide to gating in flow cytometry](https://www.bio-rad-antibodies.com/blog/a-guide-to-gating-in-flow-cytometry.html)

## Singletons vs. cell clusters
### Raw scale
```{r,fig.width=6,fig.height=7}
ggplot(dtl,
       aes(x = `FSC-A`,
           y = `SSC-A`,
           color = Group)) +
  facet_wrap(~ `File Name`) +
  geom_point(shape = ".",
             alpha = 0.3) +
  # geom_density_2d() +
  stat_density_2d(contour_var = "count",
                  color = "black") +
  theme_bw() +
  theme(legend.position = "none")
```

## Histograms
```{r}
trans2t(dtl$`FITC-A`)

tmp <- dtl[, .(`trans FITC-A` = trans2t(`FITC-A`),
               `trans PE-A` = trans2t(`PE-A`),
               `trans PerCP-Cy5-5-A` = trans2t(`PerCP-Cy5-5-A`),
               `trans PE-Cy7-A` = trans2t(`PE-Cy7-A`), 
               `trans APC-A` = trans2t(`APC-A` ),
               `trans APC-Cy7-A` = trans2t(`APC-Cy7-A`),
               `trans Pacific Blue-A` = trans2t(`Pacific Blue-A`),
               `trans Alex 700-A` = trans2t(`Pacific Blue-A`))]

ggplot(dtl,
       aes(x = log2(`FITC-A` - min(`FITC-A`) + 1))) +
         facet_wrap(~ `File Name`) +
         geom_histogram(bins = 100) +
  theme_bw()
```


# Session
```{r}
sessionInfo()
```